Select-based random access \\ to variable-byte encodings

### 2020-05-24 gradu/msc_english_contents.tex
Final fixes

Another improvement to VB encoding is to see whether the compressed numbers can be changed. This is trivial until the block length of the compressed number changes. Then an additional data structure or 
cache is needed and the supporting rank/select data structure needs to be recalculated.

TODO: a few paragraphs about haswell instructions

>>>>

The $rank$ operation can be created with a small number of instructions utilizing a modern \textsc{POPCOUNT} machine instruction \citep{Zho13}. The $select$ approach is 
known to be slower, ranging from 50 to 80 machine instructions. Intel's Haswell CPUs offer two new interesting machine instructions, \textsc{PDEP} and \textsc{TZCNT}. \textsc{PDEP} stores bits
from its first operand based on the location of 1 bits in the second operand. \textsc{TZCNT} is a machine instruction to count trailing zeros. These two can be used to create a $select_1$ query on 
a machine word with just 4 instructions \citep{Pan17a}. They have been tested with the $select$ of the SDSL-library and show an improvement of 2-15\% \citep{Pan17b}. Additionally, the 
$select$ implementation proposed in this thesis uses a trailing zero count to calculate the length of the element. Some improvement in speed may happen if the implementation is changed 
to use \textsc{TZCNT} instead.

One more interesting improvement to VB encoding is to see whether the compressed numbers can be changed. This is trivial until the block length of the compressed number changes. Then an additional data structure or 
cache is needed and the supporting rank/select data structure needs to be recalculated.
